---
title: "Reproducible Research Course Proyect 1"
author: "Ramírez López Elsa Itzel"
date: "14/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reproducible Research Course Proyect 1 

This is a submission for Reproducible Research Course Proyect 1. For more details taje a look at [here] (https://github.com/ItzelRamlo/RepData_PeerAssessment1/blob/master/README.md). Also there you can find the link to download the dataset

## Loadin the Data 
In this section you can find the code needed to  read and load the data por further analysis.

```{r}
#load package
library(ggplot2)
#read the data. Here I've write the url for my workdirectory, soy you should change it. Ypu can find the url for ypur own work directory using **getwd()**
activity <- read.csv('C:/Users/ELSA LOPEZ/Documents/R/Reproducible-Research-Course-Proyect-1/activity.csv')

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
