---
title: "Reproducible Research Course Proyect 1"
author: "Ramírez López Elsa Itzel"
date: "14/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This is a submission for Reproducible Research Course Proyect 1. For more details taje a look at [here](https://github.com/ItzelRamlo/RepData_PeerAssessment1/blob/master/README.md). Also there you can find the link to download the dataset

## Loading and processing the Data 
1. Load the data (i.e. \color{red}{\verb|read.csv()|}read.csv())

2.Process/transform the data (if necessary) into a format suitable for your analysis

```{r}
#load package
library(ggplot2)
library(dplyr)
```
Here I've write the url for my workdirectory, soy you should change it. 

You can find the url for your own work directory using the output of **getwd()** instead of what I wrote in the code.

```{r}
activity <- read.csv('C:/Users/ELSA LOPEZ/Documents/R/Reproducible-Research-Course-Proyect-1/activity.csv')
Sys.setlocale('LC_TIME', 'English')
```

```{r}
#some information/summary about our data
str(activity)
```
Here we can see the variables included in the dataset:

1.**steps**: Number of *steps* in a interval of 5 minutes 

2.**date**: The *date* which the measuremen was taken. The format for this variable is *YYYY-MM-DD*

3.**intervals**:This variable is an *identifier* for the interval in which the measurement was taken. 

```{r}
summary(activity)
```
Here we can see a little summary due to have an idea of the data.

## What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in the dataset.

1. Calculate the total number of steps taken per day

2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram 
of the total number of steps taken each day

3. Calculate and report the mean and median of the total number of steps taken per day
```{r pressure, echo=FALSE}
activity$date <- as.POSIXct(activity$date, "%Y-%m-%d")
weekday <- weekdays(activity$date)
activity <- cbind(activity,weekday)
activity_total_steps <- with(activity, aggregate(steps, by = list(date), FUN = sum, na.rm = TRUE))
names(activity_total_steps) <- c("date", "steps")
hist(activity_total_steps$steps, main = "Total number of steps taken per day", xlab = "Total steps taken per day", ylim = c(0,20), breaks = seq(0,25000, by=2500))
```

The **mean** and the **median** of our data is given by:

```{r}
mean(activity_total_steps$steps)
median(activity_total_steps$steps)
```


##What is the average daily activity pattern?

1. Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
average_daily_activity <- aggregate(activity$steps, by=list(activity$interval), FUN=mean, na.rm=TRUE)
names(average_daily_activity) <- c("interval", "mean")
plot(average_daily_activity$interval, average_daily_activity$mean, lwd = 2, xlab="Interval", ylab="Average number of steps", main="Average number of steps per intervals")

```
The 5-minute interval that, on average, contains the maximum number os steeps is given by:
```{r}
average_daily_activity[which.max(average_daily_activity$mean), ]$interval
```

##Imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as \color{red}{\verb|NA|}NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with \color{red}{\verb|NA|}NAs)

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?


A missing value is when we have no observations in our measurement, in our data set, this miising values are coded as **NA´s**

fist, how many missing values we have?

```{r}
sum(is.na(activity$steps))
```
 Ok, then, we need an strategy for these vlues and create a new variable. 

```{r}
imputed_steps <- average_daily_activity$mean[match(activity$interval, average_daily_activity$interval)] 

activity_imputed <- transform(activity, steps = ifelse(is.na(activity$steps), yes = imputed_steps, no = activity$steps))
total_steps_imputed <- aggregate(steps ~ date, activity_imputed, sum)
names(total_steps_imputed) <- c("date", "daily_steps")
```
Finally, we´ll make another histogram due to take a look about the changes in our output once we've removed the missing values. 
```{r}
hist(total_steps_imputed$daily_steps, xlab = "Total steps per day", ylim = c(0,30), main = "Total number of steps taken each day", breaks = seq(0,25000,by=2500))
```

The new **mean** and **median** of our data ys given by:

```{r}
mean(total_steps_imputed$daily_steps)
median(total_steps_imputed$daily_steps)
```
##  Are there differences in activity patterns betwen weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

2. Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
activity$date <- as.Date(strptime(activity$date, format="%Y-%m-%d"))
activity$datetype <- sapply(activity$date, function(x) {
        if (weekdays(x) == "Saturday" | weekdays(x) =="Sunday") 
                {y <- "Weekend"} else 
                {y <- "Weekday"}
                y
        })
```

```{r}
activity_by_date <- aggregate(steps~interval + datetype, activity, mean, na.rm = TRUE)
plot<- ggplot(activity_by_date, aes(x = interval , y = steps, color = datetype)) +
       geom_line() +
       labs(title = "Average daily steps by type of date", x = "Interval", y = "Average number of steps") +
       facet_wrap(~datetype, ncol = 1, nrow=2)
print(plot)
```

